The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-01 Paper.Rmd
modified

This section is attributed to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data. Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness. We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula (for details see Appendix \ref{SM-Consequence_skewness}). If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness). Figure \ref{fig_skewness_methane_maintext} showed asymmetric copula structure's effect on spatial skewness for biogeochemical methane-flux data.
The result, showing significantly different skewness of spatial mean from that of Normal surrogates, was consistent with model selection and nonparametric approach which supported lower tail dependence copula for methane-flux data.

We then showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}). Finally, we studied **(iii)** the effects of non-Normal copula on extinction risk with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
\caption[Skewness-methane-flux data]{Skewness of spatially averaged methane-flux data (red line) was compared with respect to that of 10000 Normal surrogates preserving $(A)$ Kendall correlation, and $(B)$ Spearman correlation. p-value < 0.05 in both case showed significant difference in skewness (as expected for lower tail dependent methane flux data from model selection results) from Normal surrogates. \label{fig_skewness_methane_maintext}}
\end{center}
\end{figure}

>>>>

This section is attributed to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data. Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness. Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data. To test null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure, we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same. Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R). We made a histogram of surragate skewness and found if actual skewness were within $95\%$ quantile or not. If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).

We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}). For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which inturn had higher slope than upper tail dependent Survival Clayton family (Fig. \ref{SM-fig_spTL}). Finally, we studied **(iii)** the effects of non-Normal noise copula on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}). We considered patches were arranged in a linear chain fashion and the population may disperse to its nearerst two neighbors (local dispersion) or to any patches (global dispersion). For both model with both dispersion, increasing number of habitat patches and increasing carrying capacity, separately, had decreased extinction risk of the populations which was expected (Fig. \ref{SM-fig_risk_numlocs_effect_local_disp} - Fig. \ref{SM-fig_risk_K_effect}). For Lewton-Cohen model, we found greater extinction risk if populations were affected by left tail dependence environmental noise copula than right tail dependent one. For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but after $r=0.7$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf}
\caption{Skewness of spatially averaged $(A)$ Green-spruce aphid's abundance, $(B)$ Leaf-plum curling aphid's firstflight date, $(C)$ Ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red line in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Kendall correlation). \label{fig_skewness_kendall}}
\end{center}
\end{figure}

## Application : Change of skewness

we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data. Figure \ref{fig_skewness_kendall} represents the results for each dataset with Kendall correlatiomn preserving surrogates. We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula. If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).

Figure \ref{fig_skewness_kendall}(A) showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula. In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_kendall} A). As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results. On the other hand, Figure \ref{fig_skewness_kendall}(B) demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula. As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_kendall} B) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence. Figure \ref{fig_skewness_kendall}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula. Our previous results on non-parametric analysis and model fitting also supported lower tail dependence. But, as $p>0.05$, we could not reject $H_{0}$ in that case. Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_kendall} D). As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.

We found the same result when we repeated our simulation with Spearman correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_spearman}).
